Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)

### 2020-07-01 manuscript/manuscript.Rmd
general refactoring 2

home_mean_occupancy <- occupancy_interpolated  %>% 
  filter(between(date, date_min_tallies, date_max_tallies)) %>% 
  group_by(home_code) %>% 
  summarise(mean_occupancy = mean(occupancy, na.rm = T))

@@

residents_episodes <- tline %>% 
  filter(rday==1) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, occ_bedroom_ratio, no_bedrooms)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category,
                   beds,  pct_occupancy, occ_bedroom_ratio, no_bedrooms)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

@@

home_total_confirmed <- distinct(reference_homes, home_code, home_name_clickview, 
                                 country, region_UK, staff_to_bed_ratio, carer_to_bed_ratio, 
                                 bed_tot_category, beds, pct_occupancy, no_bedrooms, occ_bedroom_ratio,
                                 younger_person, residential_general, residential_dementia,
                                 nursing_general,nursing_dementia) %>% 
  left_join(res_counts, by = "home_code") %>% 
  left_join(home_total_confirmed_CT) %>% 
  left_join(home_total_confirmed_DAT) %>% 
  mutate(CT_confirmed = tidyr::replace_na(CT_confirmed, 0),
         DAT_confirmed = tidyr::replace_na(DAT_confirmed, 0))

@@

residents_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(select(new_cases, date, home_code, 
                     new_confirmed_home, new_confirmed_hospital,
                     new_symptomatic_home, new_symptomatic_hospital,
                     new_deaths_home, new_deaths_hospital)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, 
                     staff_total, nursing_dementia, nursing_general, 
                     residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  left_join(occupancy_interpolated) %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(new_symptomatic = tidyr::replace_na(new_symptomatic_home, 0) +
                                    tidyr::replace_na(new_symptomatic_hospital, 0),
         new_confirmed = tidyr::replace_na(new_confirmed_home, 0) +
                                    tidyr::replace_na(new_confirmed_hospital, 0),
         new_deaths = tidyr::replace_na(new_deaths_home, 0) +
                                    tidyr::replace_na(new_deaths_hospital, 0)) %>% 
  mutate(I_symptomatic = total_cases_in_home(occupancy, new_symptomatic, date),
         I_confirmed = total_cases_in_home(occupancy, new_confirmed, date)) %>% 
  ungroup()

@@

```{r factors_outbreak}
home_total_confirmed$outbreak <-  as.factor(home_total_confirmed$CT_confirmed>0)
# # factors_outbreak <- glm( 
# #   outbreak ~ staff_to_bed_ratio + bed_tot_category + 
# #     pct_occupancy +occ_bedroom_ratio + region_UK , 
# #   family = binomial(), data = home_total_confirmed)
# # summary(factors_outbreak)
# explanatory = c( "bed_tot_category", "pct_occupancy", "staff_to_bed_ratio",
#                  "occ_bedroom_ratio",
#                  #"younger_person", "residential_general", "factor(residential_dementia)",
#                  # "nursing_general","nursing_dementia", 
#                  "region_UK")
# dependent = "outbreak"
# odds_outbreak <-  finalfit(home_total_confirmed,
#                            dependent, 
#                            explanatory, column = T,
#                            add_dependent_label = F)  
# knitr::kable(odds_outbreak,
#              row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
#  
# ggplot(home_total_confirmed, aes(occ_bedroom_ratio, pct_occupancy, colour = outbreak)) +
#   geom_point()
```

@@

```{r CPH_confirmed,   results="asis"}
# cph_conf <- coxph(Surv(time = start_int, time2 = stop_int +1, event = confirmed) ~ gender + ageg+home_type+user_type2+region_UK+ staff_to_bed_ratio+bed_tot_category + pct_occupancy+occ_bedroom_ratio, data = residents_overview)
#  summary(cph_conf)
# cox.zph(cph_conf)
# survfit(Surv(time = start_int, time2 = stop_int +1, event = confirmed)~factor(gender), data = residents_overview) %>% ggsurvplot(ylim=c(.85,1), conf.int = T, legend = "right")
explanatory = c("gender", "ageg", "home_type", "user_type2", "region_UK" ,  
                "staff_to_bed_ratio", "bed_tot_category", "pct_occupancy")
dependent = "Surv(time = start_int, time2 = stop_int +1, event = confirmed)"
CPH_confirmed <-  finalfit(residents_overview,
                           dependent, 
                           explanatory, column = T,
                           add_dependent_label = F)  
knitr::kable(CPH_confirmed,
             row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

@@

```{r risk_death, results="asis"}
# cph_death2 <- coxph(Surv(date1, date2, death_status) ~
#                       gender+ ageg +
#                       staff_to_bed_ratio + bed_tot_category +(region_UK)+
#                       pct_occupancy + occ_bedroom_ratio + cluster(resident_id), data = residents_episodes)
# summary(cph_death2)
# cox.zph(cph_death2)
# survfit(Surv(time = date1, time2 = date2, event = death_status)~ gender, data = residents_episodes)
#  pct_occupancy, occ_bedroom_ratio,no_bedroom 
explanatory = c("gender", "ageg", "home_type", #"user_type2", 
                "staff_to_bed_ratio", "bed_tot_category", "pct_occupancy", # "occ_bedroom_ratio",
                "covid_status", "region_UK", "cluster(resident_id)")
# explanatory_interaction = c("gender", "ageg", "user_type2",
#                 "staff_to_bed_ratio", "bed_tot_category", 
#                 "covid_status:home_type", "region_UK", "cluster(resident_id)")
dependent = "Surv(time = date1, time2 = date2, event = death_status)"
CPH_death <-  finalfit(residents_episodes, dependent, explanatory, 
                       column = T, add_dependent_label = F)  
knitr::kable(CPH_death, row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))

>>>>

home_mean_occupancy <- occupancy_interpolated  %>% 
  filter(between(date, as.Date("2020-01-01"), as.Date("2020-03-14"))) %>%
  group_by(home_code) %>% 
  summarise(mean_occupancy = mean(occupancy, na.rm = T))

@@

reference_homes <- reference_homes %>% 
  mutate(occ_bedroom_ratio_g5 = cut(occ_bedroom_ratio, seq(0.2, 1.2, .2)),
         occ_bedroom_ratio_g6 = cut(occ_bedroom_ratio, seq(0.25, 1.2, .15)),
         region_UK = as.factor(region_UK)) %>% 
  mutate(occ_bedroom_ratio_g6 = relevel(occ_bedroom_ratio_g6, '(0·7,0·85]'),
         occ_bedroom_ratio_g5 = relevel(occ_bedroom_ratio_g5, '(0·8,1]'),
         region_UK <- relevel(region_UK, "Scotland"))

residents_episodes <- tline %>% 
  filter(rday==1) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

residents_overview <- tline %>% 
  filter(rday==1) %>%
  mutate(ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  group_by(resident_id, home_code, ageg, gender, 
           admission_type2, home_type, user_type2, status) %>% 
  summarise(resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed),
            symptomatic = max(-susceptible_symptomatic+1, na.rm = T),
            confirmed = max(-susceptible_confirmed+1, na.rm = T),
            start_date = min(date, na.rm = T),
            stop_date = max(date, na.rm = T),
            start_int = as.numeric(min(date, na.rm = T) - date_min),
            stop_int = as.numeric(max(date, na.rm = T)- date_min)
    ) %>% 
  ungroup() %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed"))) %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category,
                   beds,  pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  left_join(distinct(select(incidents, resident_id, starts_with("covid_first"))))

@@

home_total_confirmed <- distinct(reference_homes, home_code, home_name_clickview, 
                                 country, region_UK, staff_to_bed_ratio, carer_to_bed_ratio, 
                                 bed_tot_category, beds, pct_occupancy, no_bedrooms, 
                                 occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6,
                                 younger_person, residential_general, residential_dementia,
                                 nursing_general,nursing_dementia) %>% 
  left_join(res_counts, by = "home_code") %>% 
  left_join(home_total_confirmed_CT) %>% 
  left_join(home_total_confirmed_DAT) %>% 
  mutate(CT_confirmed = tidyr::replace_na(CT_confirmed, 0),
         DAT_confirmed = tidyr::replace_na(DAT_confirmed, 0))

@@

residents_ct <- tibble(date = seq(date_min, date_max_tallies, 1)) %>% 
  left_join(select(new_cases, date, home_code, 
                     new_confirmed_home, new_confirmed_hospital,
                     new_symptomatic_home, new_symptomatic_hospital,
                     new_deaths_home, new_deaths_hospital)) %>% 
  filter(!is.na(home_code)) %>% 
  left_join(distinct(reference_homes, home_code, home_name_datix, postcode, 
                     staff_total, nursing_dementia, nursing_general, 
                     residential_dementia, residential_general, younger_person)) %>% 
  left_join(select(reference_geography, postcode, region_UK)) %>% 
  left_join(occupancy_interpolated) %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(new_symptomatic = tidyr::replace_na(new_symptomatic_home, 0) + 
           tidyr::replace_na(new_symptomatic_hospital, 0),
         new_confirmed = tidyr::replace_na(new_confirmed_home, 0) + 
           tidyr::replace_na(new_confirmed_hospital, 0),
         new_deaths = tidyr::replace_na(new_deaths_home, 0) # +                                    tidyr::replace_na(new_deaths_hospital, 0)
         ) %>% 
  mutate(I_symptomatic = total_cases_in_home(occupancy, new_symptomatic, date),
         I_confirmed = total_cases_in_home(occupancy, new_confirmed, date)) %>% 
  ungroup()

@@

```{r factors_outbreak}
home_total_confirmed$outbreak <-  as.factor(home_total_confirmed$CT_confirmed>0)
# # factors_outbreak <- glm( 
# #   outbreak ~ staff_to_bed_ratio + bed_tot_category + 
# #     pct_occupancy +occ_bedroom_ratio + region_UK , 
# #   family = binomial(), data = home_total_confirmed)
# # summary(factors_outbreak)
# explanatory = c( "bed_tot_category", "pct_occupancy", "staff_to_bed_ratio",
#                  "occ_bedroom_ratio",
#                  #"younger_person", "residential_general", "factor(residential_dementia)",
#                  # "nursing_general","nursing_dementia", 
#                  "region_UK")
# dependent = "outbreak"
# odds_outbreak <-  finalfit(home_total_confirmed,
#                            dependent, 
#                            explanatory, column = T,
#                            add_dependent_label = F)  
# knitr::kable(odds_outbreak,
#              row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
#  
# ggplot(home_total_confirmed, aes(occ_bedroom_ratio, pct_occupancy, colour = outbreak)) +
#   geom_point() +
#   ggsci::scale_color_lancet() +
#   xlab("Ratio mean occupancy:number bedroom") + ylab("Ratio mean occupancy:beds")
```

@@

```{r CPH_confirmed,   results="asis"}
# cph_conf <- coxph(Surv(time = start_int, time2 = stop_int +1, event = confirmed) ~ gender + ageg+home_type+user_type2+region_UK+ staff_to_bed_ratio+bed_tot_category + pct_occupancy+occ_bedroom_ratio, data = residents_overview)
#  summary(cph_conf)
# cox.zph(cph_conf)
# survfit(Surv(time = start_int, time2 = stop_int +1, event = confirmed)~factor(gender), data = residents_overview) %>% ggsurvplot(ylim=c(.85,1), conf.int = T, legend = "right")
explanatory = c("gender", "ageg", "home_type", "user_type2",  
                "staff_to_bed_ratio", "bed_tot_category", "occ_bedroom_ratio_g6", "region_UK" )
dependent = "Surv(time = start_int, time2 = stop_int +1, event = confirmed)"
CPH_confirmed <-  finalfit(residents_overview,
                           dependent, 
                           explanatory, column = T,
                           add_dependent_label = F)  
knitr::kable(CPH_confirmed,
             row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

@@

```{r risk_death, results="asis"}
# cph_death2 <- coxph(Surv(date1, date2, death_status) ~
#                       gender+ ageg +covid_status+
#                       staff_to_bed_ratio + bed_tot_category +(region_UK)+
#                       pct_occupancy + occ_bedroom_ratio + cluster(resident_id), data = residents_episodes)
# summary(cph_death2)
# cox.zph(cph_death2)
# survfit(Surv(time = date1, time2 = date2, event = death_status)~ gender, data = residents_episodes)
#  pct_occupancy, occ_bedroom_ratio,no_bedroom 
explanatory = c("gender", "ageg", "home_type", #"user_type2", 
                "staff_to_bed_ratio", "bed_tot_category", "occ_bedroom_ratio_g6",
                "covid_status", "region_UK", "cluster(resident_id)")
# explanatory_interaction = c("gender", "ageg", "user_type2",
#                 "staff_to_bed_ratio", "bed_tot_category", 
#                 "covid_status:home_type", "region_UK", "cluster(resident_id)")
dependent = "Surv(time = date1, time2 = date2, event = death_status)"
CPH_death <-  finalfit(residents_episodes, dependent, explanatory, 
                       column = T, add_dependent_label = F)  
knitr::kable(CPH_death, row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
